---
title: "Project_8_allItems"
author: "MohammedBaaoum"
date: "2022-10-23"
output: html_document
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Original dataset 
```{r}
data <- read.csv("AEMS-2019.csv")
```
#Package/libraries 
```{r, include=FALSE}
#install.packages("tidyverse")
library(tidyverse)
#install.packages("cli")
library(cli)
#install.packages("rlang")
library(rlang)
library(psych)# for factor analysis 
library(psychTools)

library(ggplot2)
library (descstat)
library(dplyr)

library(REdaS)# to perform Bartlett’s Test of Sphericity before FA 
library(GPArotation) # required by FA 
library(GGally)
library(lavaan)# used for CFA and SEM modeling 
library(lavaanPlot)
library(semPlot)# also use to build graph for CFA, and SEM
library(caret)
```

```{r}
plotTheme <- function() {
  theme(
    plot.title = element_text(size = 14, family = "sans", face = "plain", hjust = 0),
    plot.subtitle=element_text(size = 11, family = "sans", hjust = 0),
    plot.caption=element_text(size = 10, family = "sans", face = "italic", hjust = 0), 
    axis.title.x = element_text(size = 10, family = "sans", face = "plain", hjust = 1, vjust = -0.5),
    axis.title.y = element_text(size = 10, family = "sans", face = "plain", hjust = 1, vjust = 1),
    axis.text = element_text(size = 9, family = "sans", face = "plain"),
    panel.background = element_blank(),
    panel.grid.minor = element_line(colour = "gray"),
    panel.grid.major = element_line(colour = "gray"),
    axis.ticks = element_blank(),
    legend.title = element_text(size = 10, family = "sans"),
    legend.text = element_text(size = 9, family = "sans"),
    axis.line = element_blank()
  )
}
```

# Exploratory Data Analysis 
```{r}
#describe(data)
# source:
#Variable Name: Source
#1 = School Student
#2 = School Teacher
#3 = University Instructor
#4 = University Student

# to count how many responses by source type 
data%>% count(data$Source)

# to see how many respondent by group (1=school students, 2= teacher, 3=Unv. instructor, 4= Univ. students)
ggplot(data, aes(x=Source))+
  geom_bar(color="darkblue", fill="lightblue")+
  labs(
    title = "Responses Per Participant Type",
    x = "1= School Student, 2 = School Teacher, 3= University Instructor, 4= University Student",
    y = "Count",
    color = "Geschlecht", 
    
  )
table(data$Source)
```

```{r}
library(naniar)# used to replace missing values (e.g. 99,88, etc) with NAs 
# summary package for descriptive stat
#library(summarytools)
```

# High school students: selection and NA investigation 
```{r}
school_student<- data[which(data$Source == "1"),]# select school students only 
names(school_student)
summary(school_student)

#The rest of the dataset (W/O Religion variable)
school_student_rest<- select(school_student,c(1:10, 12:160))

# define NAs values for the rest of the dataset
NA_values<- c("99", "88", "9")
school_studnet_rest_NA<- school_student_rest %>%
  replace_with_na_all(condition = ~.x %in% NA_values)

# Religion variable 
school_studnt_religion<- select(school_student, 11)
#dfSummary(school_studnt_religion)
school_studnt_religion_NA<- school_studnt_religion %>% replace_with_na(replace = list(Religion = 99))
summary(school_studnt_religion_NA)


# Combine the entire dataset with NAs
school_studnet_NA<- cbind(school_studnt_religion_NA, school_studnet_rest_NA)
names(school_studnet_NA)



sapply(school_studnet_NA, function(X) sum(is.na(X)))  # count na per variable

most_missing <- select(school_studnet_NA, c(1:20)) # visualizing the NA for the variable that has missing more than 10% 
gg_miss_var(most_missing, show_pct = TRUE)
```

#Remove  variables for teachers and univ. students and only keep school studnets variables 
```{r}
names(school_studnet_NA)
school_student_Variabler<- school_studnet_NA[c(-2, -5:-6, -10:-11,-14:-16, -97:-123)]# deleted:source names, resp. ID, school type, degree year, education, relationship status, children, work experience, and other construct not related to student 

names(school_student_Variabler)

most_missing_schoolStudent <- select(school_student_Variabler, c(1:20)) # visualizing the NA for the variable that has missing more than 10% 
gg_miss_var(most_missing_schoolStudent, show_pct = TRUE)

# Remove Grade and use age as an alternative. also, remove all other NAs in countery, ethincity, etc 

school_student_cleaning1 <- school_student_Variabler[c(-6)]
names(school_student_cleaning1)

# Visulize after removing Grade 
most_missing_clean1 <- select(school_student_cleaning1, c(1:35)) # visualizing the NA for the variable that has missing more than 10% 
gg_miss_var(most_missing_clean1, show_pct = TRUE)

```

#Omit NAs from school student data
```{r}
school_student_cleaning2 <- na.omit(school_student_cleaning1)
#summary(school_student_cleaning2)

# to check which row and how many obs.  =8 for the two variables with 8
which(school_student_cleaning2$SR_Problems==8)
which(school_student_cleaning2$SB_BeMyself==8)

# remove "na = 8" from SB.bemyself, and SR.problems

schoolData<- filter(school_student_cleaning2, 
                  !SR_Problems==8, 
                  !SB_BeMyself==8)

```


```{r}
#write.csv(schoolData, "school_data.csv")
```

# revers coding for some CIO Items
```{r}

reverse_cols <- c ("CIO_SelfDepend", 
                   "CIO_SelfDependMost",
                   "CIO_OwnThing",
                   "CIO_Identity",
                   "CIO_JobBetter",
                   "CIO_Competition",
                   "CIO_BetterTense",
                   
"Gratitude_NotMuch",
"Gratitude_Time",
"Empathy_Waste",
"Empathy_Difficult",
"SR_GoalSettingHard",
"SR_GoalPlanTrouble",
"SR_Distracted",
"SR_Trouble",
"SR_Decisions",
"SR_Change",
"SR_Problems",
"SR_Focus",
"SR_Mistakes",
"SB_DifferentSchool",
"SB_NotInterested",
"SB_FeelDifferent",
"SB_DontBelong",
"SB_AcceptanceHard",
"MM_NoPurpose",
"SE_Insecure",
"SE_DontHandle",
"SE_NotCapable",
"SE_RarelyAchieve",
"SE_NewGiveUp",
"SE_AvoidDifficult",
"SE_NotTryComplicated")

schoolData[,reverse_cols] <- 5 - schoolData[,reverse_cols]
str(schoolData)
summary(schoolData)
```





#Grouping itmes to single construct based on the mean  of each row 

```{r}
#package for correlation 
#install.packages("GGally")
library(GGally)
names(schoolData)
library(corrplot)

#SB
SB_data<- select(schoolData,c(65:82))
str(SB_data)

ggcorr(SB_data, method = c("everything", "pearson"))
corPlot(SB_data, cex = 0.5)

SB_mean<- SB_data %>% as_tibble() %>% 
   mutate(mean_SB = rowMeans(across(where(is.numeric))))
SB_mean

#PS
ps_data<- select(schoolData,c(43:54))
str(ps_data)
corPlot(ps_data, cex = 0.5)
ps_mean<- ps_data %>% as_tibble() %>% 
   mutate(mean_ps = rowMeans(across(where(is.numeric))))
ps_mean

#ER
ER_data<- select(schoolData,c(94:101))
corPlot(ER_data, cex = 0.5)

ER_mean<- ER_data %>% as_tibble() %>% 
   mutate(mean_ER = rowMeans(across(where(is.numeric))))
ER_mean

#Empathy 
Empathy_data <-  select(schoolData,c(102:108))
corPlot(Empathy_data, cex = 0.5)

Empathy_mean<- Empathy_data %>% as_tibble() %>% 
   mutate(mean_Empathy = rowMeans(across(where(is.numeric))))
Empathy_mean

#RS
RS_data<- select(schoolData,c(83:87))
corPlot(RS_data, cex = 0.5)

RS_mean<- RS_data %>% as_tibble() %>% 
   mutate(mean_RS = rowMeans(across(where(is.numeric))))
RS_mean

#SR
SR_data<- select(schoolData,c(109:124))
corPlot(SR_data, cex = 0.5)

SR_mean<- SR_data %>% as_tibble() %>% 
   mutate(mean_SR = rowMeans(across(where(is.numeric))))
SR_mean

#Forg
forgive_data<- select(schoolData,c(8:16))
corPlot(forgive_data, cex = 0.5)

forgive_mean<- forgive_data %>% as_tibble() %>% 
   mutate(mean_forgive = rowMeans(across(where(is.numeric))))
forgive_mean

#MM
MM_data<- select(schoolData,c(55:64))
corPlot(MM_data, cex = 0.5)

MM_mean<- MM_data %>% as_tibble() %>% 
   mutate(mean_MM = rowMeans(across(where(is.numeric))))
MM_mean
#SE
SE_data<- select(schoolData,c(31:42))
corPlot(SE_data, cex = 0.5)

SE_mean<- SE_data %>% as_tibble() %>% 
   mutate(mean_SE = rowMeans(across(where(is.numeric))))
SE_mean

#Gratidute 
gratitude_data<- select(schoolData,c(88:93))
corPlot(gratitude_data, cex = 0.5)

gratitude_mean<- gratitude_data %>% as_tibble() %>% 
   mutate(mean_gratitude = rowMeans(across(where(is.numeric))))
gratitude_mean

#CIO
CIO_data<- select(schoolData,c(17:30))
corPlot(CIO_data, cex = 0.5)

CIO_mean<- CIO_data %>% as_tibble() %>% 
   mutate(mean_CIO = rowMeans(across(where(is.numeric))))


```

#Combined the mean into one dataset 
```{r}

mean_school.Data<- cbind(CIO_mean$mean_CIO, gratitude_mean$mean_gratitude,MM_mean$mean_MM, SE_mean$mean_SE, forgive_mean$mean_forgive, SR_mean$mean_SR, RS_mean$mean_RS,ER_mean$mean_ER,Empathy_mean$mean_Empathy, ps_mean$mean_ps, SB_mean$mean_SB)

mean_school.Data <- as.data.frame(mean_school.Data)

```

#Rename variables/construct
```{r}
names(mean_school.Data)

mean_school.Data<- rename(mean_school.Data,
       CIO= "V1",
       GRAT= "V2",
       MEANMK= "V3",
       SEFF= "V4",
       FORG= "V5",
       SREG= "V6",
       RELSPIR = "V7",
       EMOREG= "V8", 
       EMP= "V9",
       PSOLV= "V10",
       SOB="V11") 
names(mean_school.Data)


```
 
# Final data for modeling 
```{r}

#add deographic data 
names(schoolData)
Demo_schoo.data<- select(schoolData, c(3:5))
names(Demo_schoo.data)
 
school <- cbind (mean_school.Data, Demo_schoo.data)# the entire data set (including demographic and construct)
str(school)
```

#Remove participants older than 24 years 
```{r}
names(school)


school_less24<- subset.data.frame(school, school$Age <= 2)
table(school_less24$Country)
names(school_less24)

# Rename Country to Region
school_rename<- rename(school_less24,
       Region= "Country")
names(school_rename)

```


# Recoding region for demographic data in the entire data set
```{r}
library(tidyverse)
school_recoding<- school_rename %>%
                   mutate(Region = recode(Region, 
                                          "15" = 1,
                                          "1" = 1,
                                          "14" = 2,
                                          "2"= 2,
                                          "4"= 3,
                                          "7" = 3, 
                                          "5" = 4,
                                          "13" = 4,
                                          "6" = 5,
                                          "11"= 5,
                                          "12" = 5, 
                                          "8" = 0,
                                          "10"= 0,))
names(school_recoding)
table(school_recoding$Region)


```


# Remove demographic varibales as suggested by stakeholder 
```{r}


school_model<- (school_recoding)
names(school_model)
str(school_model)
```
#Variable visulization after taking mean  
```{r}
str(school_model)
par(mfrow=c(2,2))
hist(school_model$CIO)
hist(school_model$GRAT)
hist(school_model$MEANMK)
hist(school_model$SEFF)
hist(school_model$FORG)
hist(school_model$SREG)
hist(school_model$RELSPIR)
hist(school_model$EMOREG)
hist(school_model$EMP)
hist(school_model$PSOLV)
hist(school_model$SOB)


# Summary statistics for all variables table 
#Source:https://cran.r-project.org/web/packages/vtable/vignettes/sumtable.html
library(vtable)
#st(school_model)

sumtable(school_model,vars = c('CIO','Grat','MM','Self.Efc','Self_Reg', 'Relg.Spit','ER','EMP','PR.S', 'Self.Bel', 'Forg'),
         summ=c('notNA(x)',
                'mean(x)',
                'median(x)',
                'sd(x)',
                'min(x)',
                'max(x)'))
table(school_model$Region)


```


# Use Cronbach’s Alpha to check the the items for each factor
```{r}

forgive<- schoolData[, c("Forgive_BrokenEngagement",
                                      "Forgive_SecretDisclosure",
                                      "Forgive_CousinArgument",
                                      "Forgive_CurseSameReligion",
                                      "Forgive_CurseDiffReligion",
                                      "Forgive_Wall",
                                      "Forgive_Rumor",
                                      "Forgive_Loss", 
                                      "Forgive_Car")]

CIO<- schoolData[, c( "CIO_PeerPrize",
                                   "CIO_PeerWellbeing",
                                   "CIO_Cooperate",
                                   "CIO_Family",
                                   "CIO_RespectGrpDecisions",
                                   "CIO_SelfDepend",
                                   "CIO_SelfDependMost",
                                    "CIO_OwnThing",
                                   "CIO_Identity",
                                   "CIO_JobBetter" ,
                                    "CIO_Competition", 
                                   "CIO_BetterTense",
                                   "CIO_PleasureTime", 
                                   "CIO_ParentsChildren")]

SE<- schoolData[, c("SE_NotTryComplicated",
                                 "SE_AvoidDifficult",
                                 "SE_NewGiveUp",
                                 "SE_NotCapable",
                                 "SE_DontHandle",
                                 "SE_PlansCertain",
                                 "SE_KeepTrying",
                                 "SE_Unpleasant",
                                 "SE_RightToWork",
                                 "SE_TryHarder",
                                 "SE_RarelyAchieve",
                                 "SE_Insecure")]

PS<- schoolData[, c("PS_ResultsThink",
                                 "PS_GatherInfo",
                                 "PS_IdentifyOptions",
                                 "PS_ExpressThoughts",
                                 "PS_GiveReasons",
                                 "PS_InfoToSupport",
                                 "PS_MoreThanOne",
                                 "PS_PlanInfo",
                                 "PS_SupportDecisions",
                                 "PS_CompareIdeas",
                                 "PS_MindOpen",
                                 "PS_ListenIdeas")]

MM<- schoolData[, c("MM_LifeMeaningful",
                                              "MM_LifesPurpose",
                                              "MM_FeelSignificant",
                                              "MM_Mission",
                                              "MM_LifeMeaning",
                                              "MM_UnderstandLife",
                                              "MM_Purpose",
                                              "MM_LifeMeaningfulSense",
                                              "MM_SatisfyingPurpose",
                                                  "MM_NoPurpose")]
                  

# #"SB_DontBelong",
SB<- schoolData[, c("SB_PartOfCommunity",
                                 "SB_TeachersRespect",
                                 "SB_TreatedRespect",
                                "SB_AcceptanceHard",
                                 "SB_NoticeGood",
                                 "SB_ProudSchool",
                                 "SB_LikeMe",
                                 "SB_OpinionsSeriously",
                                 "SB_TeachersInterested",
                                 "SB_CanTalk",
                                 "SB_PeopleFriendly",
                                 "SB_Activities",
                                 "SB_BeMyself",
                                 "SB_GoodWork",
                                 "SB_DifferentSchool",
                                "SB_DontBelong",
                                "SB_FeelDifferent",
                                "SB_NotInterested")]

RS<- schoolData[, c("RS_Religion",
                                 "RS_Prayer",
                                 "RS_FeelGod",
                                 "RS_ReligionGrp",
                                 "RS_DefiningYou")]


#""Gratitude_Time"was removed and alpha increased to 0.57
Gratitude<- schoolData[, c( "Gratitude_Thankful",
                                         "Gratitude_LongList",
                                         "Gratitude_NotMuch",
                                         "Gratitude_Variety" ,
                                         "Gratitude_Appreciate"
                                         )]



 ER <- schoolData[, c("ER_NegativeThink",
                      "ER_PositiveThink",
                      "ER_ExpressPositive",
                      "ER_StressCalm",
                      "ER_PositiveChange",
                      "ER_EmotionControl",
                      "ER_LessNegative",
                      "ER_NegativeExpress")]
                                          
#"Empathy_Waste" was removed and alpha increased to 60
 Empth <- schoolData[, c("Empathy_Perspective",
                        "Empathy_TwoSides",
                        "Empathy_Upset",
                        "Empathy_Criticizing",
                        "Empathy_Sides",
                        "Empathy_Difficult")]
 
 
                                                      

SR<- schoolData[, c(  "SR_GoalSettingHard",
                      "SR_GoalPlanTrouble", 
                      "SR_Distracted",
                      "SR_Trouble", 
                      "SR_Decisions", 
                      "SR_Change",
                      "SR_Problems",
                      "SR_Focus", 
                      "SR_Mistakes",
                      "SR_Goals",
                      "SR_GoalPlan",
                      "SR_Resolution",
                      "SR_GoalProgress",
                      "SR_Willlpower",
                      "SR_MistakeOnce",
                      "SR_MistakesLearn")]

                  
                 

# Calculate Cronbach’s Alpha 
#par(mfrow=c(3,4))
alpha(forgive, check.keys = TRUE)$total[1]# 
alpha(CIO, check.keys = TRUE)$total[1] # 
alpha(SE, check.keys = TRUE)$total[1]# 
alpha(PS, check.keys = TRUE)$total[1]# 
alpha(MM, check.keys = TRUE)$total[1] #
alpha(SB, check.keys = TRUE)$total[1]#
alpha(RS, check.keys = TRUE)$total[1]#
alpha(Gratitude, check.keys = TRUE)$total[1] # 
alpha(ER, check.keys = TRUE)$total[1]#
alpha(Empth, check.keys = TRUE)$total[1]# 
alpha(SR, check.keys = TRUE)$total[1]# 

#CIO HAD 0.24
#alpha(CIO, check.keys = TRUE)


```


```{r}
library(ggplot2)# for correlqtion plot 
library(ggExtra)
library(corrplot)
# Corrplot source: https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html
```

#School_Continuse variabels/construct: this is to check the construct correlation 
```{r}
school_continuse_Variable<- select(school_model,1:11 )
names(school_continuse_Variable)

# Visualization fo the correlation 
corPlot(school_continuse_Variable, cex = 0.8)
pairs.panels(school_continuse_Variable, pch = 20)
correlation_for_factor<- round(cor(school_continuse_Variable), 2)
#write.csv(correlation_for_factor, "correlation matrix for factors.csv")

#Source:https://r-coder.com/correlation-plot-r/#:~:text=Correlation%20plots%2C%20also%20known%20as,with%20different%20functions%20and%20packages.


```
# ploting 
```{r}
boxplot(school_continuse_Variable)
str(school_continuse_Variable)

# par(mfrow=c(2,2), mai=c(.6,.5,.2,.1))
# for(i in c(2:18)){
# plot(school_continuse_Variable [,i],school_continuse_Variable$Self.Efc, xlab=names(school_continuse_Variable)[i],ylab="Self-Efficacy", col=3, pch=20)
#   
#  plotting<-  school_continuse_Variable %>%
#   ggplot( aes(x=school_continuse_Variable$Grat, y=school_continuse_Variable$Self.Efc, fill=school_continuse_Variable)) +
#     geom_violin() +
#     scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
#     theme_ipsum() +
#     theme(
#       legend.position="none",
#       plot.title = element_text(size=11)
#     ) +
#     ggtitle("Violin chart") +
#     xlab("")
#  plotting
# }
```









# Replaicng index in the data with text for demographic visulization 
```{r}

#Demo_schoo.data2 <- recode_factor(Demo_schoo.data, '0' = "Female", '1' = "male")
library(forcats)
library(tidyverse)
library(dplyr)
library(plyr)
library(car)



names(school_model)
demographic_data<- select(school_model, 12:13)
names(demographic_data)

dempgraphic_std<- demographic_data %>%
                   mutate(Gender = recode_factor(Gender, 
                                         "0" = "Female", 
                                         "1" = "Male"))


# #dempgraphic_std<- dempgraphic_std %>% # age was eleiminated from the data after only inclduing 18 or below
#                    mutate(Age = recode_factor(Age, 
#                                        "1" = "Less than 18", 
#                                        "2" = " 18_24"))
                                       
dempgraphic_std<- dempgraphic_std %>%
                   mutate(Region = recode_factor(Region,
                                          "1"=	"Ungrouped",
                                          "2"	= "Southeast Asia",
                                          "3"	= "Central Asia",
                                          "4"= "MENA",
                                          "5"= "Sub-Saharan Africa",
                                          "0"= "South Asia"))




plot(dempgraphic_std$Gender)
gender.count<-table(dempgraphic_std$Gender)



```


#Visulaization for Dempgraphic Variables

```{r}
library(ggplot2)
theme_set(theme_classic())

# Histogram on a Categorical variable
g1 <- ggplot(dempgraphic_std, aes(Region))
g1 + geom_bar(aes(fill=Gender), width = 0.5) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Gender Across Region")


#Gender only visulization
# Histogram on a Categorical variable

ggplot(data = dempgraphic_std,  aes(x = Gender))+
  geom_bar(color = 'red')


```

# Plotting IVs against DV (self-efficacy)
```{r}
#####
par(mfrow=c(3,4))
school_model %>% 
  ggplot(
    aes(x = CIO, 
        y= Self.Efc))+
  geom_point()+
  stat_smooth(se = 0, method = 'lm')
school_model %>% 
  ggplot(
    aes(x = Grat, 
        y= Self.Efc))+
  geom_point()+
  stat_smooth(se = 0, method = 'lm')
school_model %>% 
  ggplot(
    aes(x = MM, 
        y= Self.Efc))+
  geom_point()+
  stat_smooth(se = 0, method = 'lm')   
school_model %>% 
  ggplot(
    aes(x = Self.Bel, 
        y= Self.Efc))+
  geom_point()+
  stat_smooth(se = 0, method = 'lm')
school_model %>% 
  ggplot(
    aes(x = Forg, 
        y= Self.Efc))+
  geom_point()+
  stat_smooth(se = 0, method = 'lm')
school_model %>% 
  ggplot(
    aes(x = Self_Reg, 
        y= Self.Efc))+
  geom_point()+
  stat_smooth(se = 0, method = 'lm')
school_model %>% 
  ggplot(
    aes(x = ER, 
        y= Self.Efc))+
  geom_point()+
  stat_smooth(se = 0, method = 'lm')
school_model %>% 
  ggplot(
    aes(x = EMP, 
        y= Self.Efc))+
  geom_point()+
  stat_smooth(se = 0, method = 'lm')
school_model %>% 
  ggplot(
    aes(x = PR.S, 
        y= Self.Efc))+
  geom_point()+
  stat_smooth(se = 0, method = 'lm')
```


```{r, include=FALSE}
library(mlbench)
library(caret)# an aggregator package for performing many machine learning models
library(e1071)
library(lime)

library(rsample)      # data splitting 

library(ranger)       # a faster implementation of randomForest
library(h2o)          # an extremely fast java-based platform
library(rsq)
```
# Further filtiring the data based on age (less than 18 yrs only)
```{r}
names(school_model)
str(school_model)
head(school_model)
table(school_model$Age)
# filter datset based on age less than 18 yrs
#school_model<- school_model[-c(14,7)]# remove age and religious/spirtuality
school_model<- school_model[which(school_model$Age == "1"),]

# remove age column 
school_model<- school_model[-c(14)]# remove age 
 

school_model$Region<- as.factor(school_model$Region)
school_model$Gender<- as.factor(school_model$Gender)
```


# Split data for regression 
```{r}
school_modelReg<- school_model# to make the name of the data sepcific for regression 
str(school_modelReg)

set.seed(12346)
ind <- sample(2, nrow(school_modelReg), replace = T, prob = c(0.8, 0.2))
train_reg <- school_modelReg[ind == 1,]
test_reg <- school_modelReg[ind == 2,]


```

#Package used for regression 
```{r, include=FALSE}
library(randomForest)

library(DAAG)
library(MASS)
library(compositions)

library(car)
library(PerformanceAnalytics)
library(olsrr)# to run all possible soltuions for regression: forward, backword, both
library(performance)# to cehkc linear regression assumption 
library(glmnet)# for lasso regression 
install.packages("glmnet", repos = "http://cran.us.r-project.org")
```

# Regression model 
```{r}

str(train)
set.seed(1234)
reg.1<- lm(Self.Efc~., data = school_model)
summary(reg.1)
check_model(reg.1)# to cehck model assumption 

#try all possible steps: this help to try all combination of varaiables and choose best model based on R-sq, R-adjusted, CP
(all<- ols_step_all_possible(reg.1))
#write.csv(all, "model evalution_with_school_model_data.csv")
plot(all)


k <- ols_step_best_subset(all)

###Partial R-Sequard 
rsq.partial(reg.1)

# Try with step AIC: the lower AIC is the better 
model<- stepAIC(reg.1, direction = "both", trace = F)# both= forward and backword & can be changed "forward" or "backword"
summary(model)

# Prediction of the mode 
p<- predict(reg.1, test_reg)

###########Regtression without religion#######################

reg.2<- lm(Self.Efc~.-Relg.Spit, data = train_reg)
summary(reg.2)


```

```{r}
# to run modelLookUp function to check which paramneter a model has
install.packages(c("pkg1", "pkg2"), dependencies = TRUE)
library(pkg)
```

# Split data to training and testing and select variables significant based on regression 
```{r}
set.seed(1234)
ind <- sample(2, nrow(school_model), replace = T, prob = c(0.8, 0.2))
train <- school_model[ind == 1,]
test <- school_model[ind == 2,]

#names(school_model)
#write.csv(school_model, "school_model_afterVARNameChange.csv")
#write.csv(school_model, "school_model.csv")
```


# cv SET UP 
```{r}
# Regression Tree
# Bagging
set.seed(1234)
cvcontrol <- trainControl(method="cv", #here used CV (used with regression or classification)
                          number = 10,
                          allowParallel=TRUE)

```

```{r}
#Lasso  regression 
#sources:
#https://www.datacareer.ch/blog/ridge-and-lasso-in-r/
#http://www.sthda.com/english/articles/37-model-selection-essentials-in-r/153-penalized-regression-essentials-ridge-lasso-elastic-net/
#https://bookdown.org/tpinto_home/Regularisation/lasso-regression.html
#https://github.com/tirthajyoti/R-stats-machine-learning/blob/master/Stepwise%20regression%2C%20LASSO%2C%20Elastic%20Net.R
#https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r

# Spliting training set into two parts based on outcome: 75% and 25%
# Split data into train and test sets
ind <- sample(2, nrow(school_model), replace = T, prob = c(0.8, 0.2))
x.train <- school_model[ind==1, ]
x.test <- school_model[ind==2, ]

y.train <- school_model[ind==1,]
y.test <- school_model[ind==2,]





train_rows <- sample(1:n, n/2)
x.train <- x[train_rows, ]
x.test <- x[-train_rows, ]

y.train <- y[train_rows]
y.test <- y[-train_rows]
#######

cv_lasso<- cv.glmnet(train, train$Self.Efc, alpha = 1)
lasso2<- glmnet()
set.seed(1234)
lasso <- train(Self.Efc ~., 
             data=train,
             method="lasso",
             trControl=cvcontrol)

plot(lasso)

plot(varImp(lasso))
summary(lasso)

lasso$

# Plot, RMSE, R-square
ba <- predict(lasso,  test)# now do test/predict
plot(ba ~ test$Self.Efc, main = 'Predicted Vs Actual MEDV - Test data')
sqrt(mean((test$Self.Efc - lasso)^2))
cor(test$Self.Efc, lasso) ^2

str(test)


#####based on class multivariate ###########################3
#column 8 is the response variable
#source: https://rstudio-pubs-static.s3.amazonaws.com/511869_1ac4119203c6450b8e97c5058a52fb0c.html#lasso-regression
library(glmnet)# for lasso regression 
#install.packages("glmnet", repos = "http://cran.us.r-project.org")
any(is.na(school_model))
names(school_model)
y<-as.data.frame(school_model[,4]) 
x<-as.data.frame(school_model[,-4])





## Last attempt 
str(train)
set.seed(1234)
tune_lass<- expand.grid(alpha = 1,
                        lambda = seq(0.0001, 1, length = 5))
set.seed(1234)
lasso<- train(Self.Efc~., 
              train,
              method = 'glmnet',
              tuneGrid = tune_lass,
              trControl = cvcontrol)
plot(lasso)
lasso# i still have some NAs in the resutls, there are some error i nthe mdoel 
plot(lasso$finalModel, xvar = 'lambda', label = T)
plot(lasso$finalModel, xvar = 'dev', label = T)
plot(varImp(lasso))
####
```


#DT-regession using bagging 
```{r}
#-Gender -Relg.Spit -EMP 
set.seed(1234)
bag <- train(SEFF ~ ., 
             data=train,
             method="treebag",
             trControl=cvcontrol,
             importance=TRUE)# this to bring important variabeles 

plot(varImp(bag))
# install.packages("ISLR")
# library(ISLR)
# install.packages("rpart")
# library(rpart)



# Plot, RMSE, R-square
set.seed(1234)
ba <- predict(bag,  test)# now do test/predict
plot(ba ~ test$SEFF, main = 'Predicted Vs Actual MEDV - Test data')
sqrt(mean((test$SEFF - ba)^2))
cor(test$SEFF, ba) ^2


```


```{r}
##Part 2: Random forest 
# RF: same as DT except change method =RF #-Gender -Relg.Spit -EMP
set.seed(1234)
forest <- train(SEFF ~ .,  
                data=train,
                method="rf", # change the method to RF
                trControl=cvcontrol,
                tuneLength= 12,
                importance=TRUE)
# Plot mtry and find the best value
plot(forest)
forest$bestTune

# re-run the model using (mtry=2)
set.seed(1234)
forest_mtry <- train(SEFF ~ .,  
                data=train,
                method="rf", # change the method to RF
                trControl=cvcontrol,
                tuneLength= 3,
                importance=TRUE)

# plot important variabels 
plot(varImp(forest_mtry))




# Plot, RMSE, R-square based on test data 
set.seed(1234)
rf <-  predict(forest_mtry,  test)
plot(rf ~ test$SEFF, main = 'Predicted Vs Actual MEDV - Test data')
sqrt(mean((test$SEFF - rf)^2))
cor(test$SEFF, rf) ^2

head(rf)
head(school_modelReg$Self.Efc)
# Explain predictions: ML is black box and this give more description about inside
explainer <- lime(test[1:3,], forest, n_bins = 5)
explanation <- explain( x = test[1:3,], 
                        explainer = explainer, 
                        n_features = 5)
plot_features(explanation)
plot_explanations(explanation)


install.packages("rattle")
library(rattle)
fancyRpartPlot(forest$finalModel)

plot(forest$finalModel)

###Random forest based on class note############

library(randomForest)
set.seed(1)
bag.boston=randomForest(Self.Efc~.,data=train,mtry=9,trControl=cvcontrol,
importance=TRUE)
```

# Part 3: xgbtree

```{r}
# Boosting#-Gender -Relg.Spit -EMP
set.seed(1234)
boo <- train(SEFF ~ ., 
             data=train,
             method="xgbTree", 
             trControl=cvcontrol,
             tuneGrid = expand.grid(nrounds = 500,# here is the difference, xgbtree has 7 parameters (play with max_depth, eta, gamma)
                                    max_depth = 6,# this is how long the tree is (deafult =6)
                                    eta = 0.2,# learning rate (deafult =0.3)
                                    gamma = 2.1,# used to avoiding overfit (deafult =0)
                                    colsample_bytree = 1,
                                    min_child_weight = 1,
                                    subsample = 1))
plot(varImp(boo))

# Plot, RMSE, R-square
bo <-  predict(boo,  test)
plot(bo ~ test$SEFF, main = 'Predicted Vs Actual MEDV - Test data')
sqrt(mean((test$SEFF - bo)^2))
cor(test$SEFF, bo) ^2

```
```{r}
library(nnet)
library(devtools)

```

```{r}
#Source: https://beckmw.wordpress.com/2013/11/14/visualizing-neural-networks-in-r-update/
library(devtools)
library(nnet)
install.packages("NeuralNetTools")

library(NeuralNetTools)

# Neural network# -Gender -Relg.Spit -EMP
set.seed(1234)
NN <- train(SEFF ~ ., 
             data=train,
             method="nnet", 
             trControl=cvcontrol,
             linout=TRUE,
             trace = FALSE)


plot(varImp(NN))

plot(NN$finalModel)
plot.nnet(NN)
plotnet(NN)


# Plot, RMSE, R-square
NN_pred <-  predict(NN,  test)
#plot(NN ~ test$Self.Efc, main = 'Predicted Vs Actual MEDV - Test data')
sqrt(mean((test$SEFF - NN_pred)^2))
cor(test$SEFF, NN_pred) ^2

```

# using VIP fucntion to combine all importance variables 
```{r}
library(vip)# to visulzie important variabels from different models 
library(gridExtra)
```

```{r}

# Construct ggplot2-based VIPs for each model 
p1 <- vip(bag)  # CART-like decision tree
p2 <- vip(forest, width = 0.5, aesthetics = list(fill = "green3"))   # RF
#> Warning in vip.default(rfo, width = 0.5, aesthetics = list(fill = "green3")):
#> Arguments `width`, `alpha`, `color`, `fill`, `size`, and `shape` have all been
#> deprecated in favor of the new `mapping` and `aesthetics` arguments. They will
#> be removed in version 0.3.0.
p3 <- vip(boo, aesthetics = list(col = "purple2"))   # GBM
p4 <- vip(NN, aesthetics = list(col = "purple3"))   # GBM


# Display all three plots side by side
grid.arrange(p1, p2, p3,p4, ncol = 2, theme = plotTheme())

grid.arrange(p1, p2, p3, p4, ncol = 2, theme = plotTheme())

p1 <- p1+plotTheme()
p2 <- p2+plotTheme()
p3 <- p3+plotTheme()
p4 <- p4+plotTheme()
# Arrange the plots using grid.arrange
grid.arrange(p1, p2, p3, p4, ncol = 2)
```

```{r}
p1 <- vip(bag)+ ggtitle("Bagging") + theme(plot.title = element_text(hjust = 0.5))  
p2 <- vip(forest)+ ggtitle("Random Forest") + theme(plot.title = element_text(hjust = 0.5)) 
p3 <- vip(boo)+ ggtitle("XGBoost") + theme(plot.title = element_text(hjust = 0.5)) 
p4 <- vip(NN)+ ggtitle("Neural Network") + theme(plot.title = element_text(hjust = 0.5)) 

p1 <- p1+plotTheme()
p2 <- p2+plotTheme()
p3 <- p3+plotTheme()
p4 <- p4+plotTheme()
# Arrange the plots using grid.arrange
All_model <- grid.arrange(p1, p2, p3, p4, ncol = 2)

ggsave("All_model.png", All_model)
ggsave("p2.png", p2)

```


```{r}
library(reshape2)
names(school_model)
school_model_corr <- school_model %>% 
  select(c(-12, -13))

# Compute the correlation matrix
corr_matrix <- round(cor(school_model_corr),2)

# Reshape the correlation matrix to long format
corr_data2 <- melt(corr_matrix)
head(corr_data2)

corr_data <- melt(corr_matrix_filtered) %>%
  rename(Var1 = Var1, Var2 = Var2, Correlation = value) 

# Create the correlation plot
correlation_plot <- ggplot(data = corr_data, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                       midpoint = 0, limits = c(-1, 1), na.value = "gray", name = "Correlation") +
  labs(title = "Pearson Correlation Matrix") +
  theme_minimal()+
  plotTheme()+
  theme(# to remove Var 1 and Var 2 from the plot on x & y
  axis.title.x = element_blank(),
  axis.title.y = element_blank())

# Display the correlation plot
print(correlation_plot)

ggsave("correlation_plot.png", correlation_plot, , bg = "white")
```

# to plot models performance 
```{r}
# Create a data frame with model names and performance metrics
model_names <- c("Bagging", "Random Forest", "XGBoost", "Neural Network")
r_squared_train <- c(0.3066, 0.3861, 0.3501, 0.3622)
rmse_train <- c(0.3411, 0.3229, 0.3307, 0.3267)
r_squared_test <- c(0.3107, 0.3891, 0.3661, 0.3906)
rmse_test <- c(0.3446, 0.3267, 0.3317, 0.3231)

# Create a data frame for the model performance metrics
performance_df <- data.frame(Model = rep(model_names, 2),
                             Metric = rep(c("R_Squared", "RMSE"), each = length(model_names)),
                             Value = c(r_squared_train, rmse_train, r_squared_test, rmse_test),
                             Dataset = rep(c("Training", "Testing"), each = length(model_names)))

# Plot comparing model performance
performance_plot <- ggplot(performance_df, aes(x = Model, y = Value, color = Dataset, group = Dataset, linetype = Metric)) +
  geom_line() +
  labs(title = "Model Performance Comparison", y = "Metric Value") +
  scale_color_manual(values = c("Training" = "blue", "Testing" = "red")) +
  scale_linetype_manual(values = c("R_Squared" = "solid", "RMSE" = "dashed")) +
  theme_minimal() +
  theme(legend.title = element_blank())

# Display the performance plot
print(performance_plot)
```


